\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2020-01-16 gwnnDraft02.tex
non-runaway material

\documentclass[nofootinbib,twocolumn,preprintnumbers]{revtex4-1}
%\usepackage[dvips]{graphicx}
\pdfoutput=1
\usepackage{amsmath,amsthm,amssymb,multirow,psfrag}
\usepackage{epsfig}
\usepackage{color}
\usepackage{slashed}
\usepackage{soul}
\graphicspath{{./Figures/}}

@@

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of $N$ hidden sectors that undergo a phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{\textrm{GW}}^{*}(f)$, will redshift to today's value $\Omega_{\textrm{GW}}^{0}(f) = \mathcal{A}\, \Omega_{\textrm{GW}}^{*}((a_{0}/a) f) $. The redshifting factor, $\mathcal{A}$, accounts for the redshifting of both $\rho_{\textrm{GW}} $ and $\rho_{c}$~\cite{Breitbach:2018ddu,PhysRevD.49.2837},
\begin{equation}\label{eqn::redshifting}
\mathcal{A} \equiv \bigg( \frac{a}{a_{0}} \bigg)^4 \bigg( \frac{H}{H_{0}}\bigg)^2
\end{equation}
where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,
\begin{equation}
\Omega_{\textrm{GW}} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{\textrm{GW}}^{i,*}((a_{0}/a)_{i} f) .
\end{equation}
We assume that the parameters of the SFOPT do not differ between sectors: the relativistic degrees of freedom, phase transition rate, and the dark QCD scale, are all similar. This makes the redshifting factor $\mathcal{A}^i$ independent of sector number.
\begin{figure}[h!]
\centering
\includegraphics[scale=.3]{Nnatral.png} 
\caption{ Gravitational wave spectral energy density (solid curves) for standard $N$naturalness using the scalar reheaton model of section~\ref{sec:reheat}. The curve corresponding to the sum of the sectors is approximately equal to the $i=1$ curve. All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$.  The shaded dashed curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity as described in Section~\ref{sec:detection}. The ones on the right are space-based interferometers: Lisa~\citep{Audley:2017drz} (blue), DECIGO~\citep{10.1093/ptep/pty078} (light blue), BBO~\citep{PhysRevD.72.083005} (red). The ones on the left are for the pulsar timing array SKA~\cite{Janssen:2014dka} for exposure time of 5-years (purple), 10-years (orange), and 20-years (green). }
\label{fig::Nnatural}
\end{figure}
Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in Fig.~\ref{fig::Nnatural}. Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, and the construction of these curves is detailed in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sector's contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

@@

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=.45\textwidth ]{highest.png}
\hfill
\includegraphics[width=.45\textwidth]{TwoFar.png} 
\hfill
\includegraphics[width=.45\textwidth]{TwoMed.png} 
\hfill
\includegraphics[width=.45\textwidth]{energydensity.png} 
\end{minipage}
\hfill
\caption{ Gravitational wave spectral energy density for the various scenarios found in Tab.~(\ref{tab:cases}).  All contributions are assumed to be purely from runaway bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$. The inset is a closer look at the region around the peaks. The shaded curves are the same as Fig.~\ref{fig::Nnatural}.   }
\label{fig:Haa}
\end{figure*}

@@

>>>>

\documentclass[nofootinbib,twocolumn,preprintnumbers]{revtex4-1}
%\usepackage[dvips]{graphicx}
\pdfoutput=1
\usepackage{amsmath,amsthm,amssymb,multirow,psfrag}
\usepackage{epsfig}
\usepackage{color}
\usepackage{slashed}
\usepackage{soul}
\usepackage{appendix}
\graphicspath{{./Figures/}}

@@

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of $N$ hidden sectors that undergo a phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{\textrm{GW}}^{*}(f)$, will redshift to today's value $\Omega_{\textrm{GW}}^{0}(f) = \mathcal{A}\, \Omega_{\textrm{GW}}^{*}((a_{0}/a) f) $. The redshifting factor, $\mathcal{A}$, accounts for the redshifting of both $\rho_{\textrm{GW}} $ and $\rho_{c}$~\cite{Breitbach:2018ddu,PhysRevD.49.2837},
\begin{equation}\label{eqn::redshifting}
\mathcal{A} \equiv \bigg( \frac{a}{a_{0}} \bigg)^4 \bigg( \frac{H}{H_{0}}\bigg)^2
\end{equation}
where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,
\begin{equation}
\Omega_{\textrm{GW}} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{\textrm{GW}}^{i,*}((a_{0}/a)_{i} f) .
\end{equation}
We assume that the parameters of the SFOPT do not differ between sectors: the relativistic degrees of freedom, phase transition rate, and the dark QCD scale, are all similar. This makes the redshifting factor $\mathcal{A}^i$ independent of sector number.
\begin{figure}[h!]
\centering
\includegraphics[scale=.3]{Nnatral.png} 
\caption{ Gravitational wave spectral energy density (solid curves) for standard $N$naturalness using the scalar reheaton model of section~\ref{sec:reheat}. The curve corresponding to the sum of the sectors is approximately equal to the $i=1$ curve. All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$. The colored solid lines use $\beta/H  = 10$ where as the dashed gray line is the total contribution of all sectors for $\beta/H  = 10^4$.  The shaded dashed curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity as described in Section~\ref{sec:detection}. The ones on the right are space-based interferometers: Lisa~\citep{Audley:2017drz} (blue), DECIGO~\citep{10.1093/ptep/pty078} (light blue), BBO~\citep{PhysRevD.72.083005} (red). The ones on the left are for the pulsar timing array SKA~\cite{Janssen:2014dka} for exposure time of 5-years (purple), 10-years (orange), and 20-years (green). }
\label{fig::Nnatural}
\end{figure}
Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in Fig.~\ref{fig::Nnatural}. Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, and the construction of these curves is detailed in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sector's contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

@@

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=.45\textwidth ]{highest.png}
\hfill
\includegraphics[width=.45\textwidth]{TwoFar.png} 
\hfill
\includegraphics[width=.45\textwidth]{TwoMed.png} 
\hfill
\includegraphics[width=.45\textwidth]{energydensity.png} 
\end{minipage}
\hfill
\caption{ Gravitational wave spectral energy density for the various scenarios found in Tab.~(\ref{tab:cases}).  All contributions are assumed to be purely from runaway bubble collisions $\Omega_{\phi}$. The colored solid lines use $\beta/H  = 10$ where as the dashed gray line is the total contribution of all sectors for $\beta/H  = 10^4$. The inset is a closer look at the region around the peaks for the $\beta/H  = 10$ case. The shaded curves are the same as Fig.~\ref{fig::Nnatural}.   }
\label{fig:Haa}
\end{figure*}

@@

\appendix
\section*{Appendix: Non-runaway phase transitions}
\label{sec:appendix}

If the phase transition occurs in the non-runaway regime $(\alpha_{\infty} > \alpha)$, the dominant contributions to the GW energy densities are given by the sound wave $h^2\Omega_{v}$ and MHD $h^2\Omega_{v}$ components,
\begin{equation}\label{eqn:Spectralnonrun}
h^2\Omega_{\textrm{GW}} \approx  h^2\Omega_{v} + h^2\Omega_{turb}.
\end{equation}

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=.45\textwidth ]{MaxSW.png}
\hfill
\includegraphics[width=.45\textwidth]{LargeSW.png} 
\hfill
\includegraphics[width=.45\textwidth]{MedSW.png} 
\hfill
\includegraphics[width=.45\textwidth]{FiveSW.png} 
\end{minipage}
\hfill
\caption{ Gravitational wave spectral energy density for the various scenarios found in Tab.~(\ref{tab:cases}).  All contributions are assumed to be from a non-runaway phase transition with terminal velocity $v = 0.95$. The colored solid lines use $\beta/H  = 10$ where as the dashed gray line is the total contribution of all sectors for $\beta/H  = 10^4$ . The shaded curves are the same as Fig.~\ref{fig::Nnatural}. In contrast to the runaway case, most scenarios evade the projected sensistivies of future experiments and therefore is the less optimal situation }
\label{fig:Haasw}
\end{figure*}

The new contributions to the GW energy density take on a different form from Eq.~(\ref{eqn::GWengdensity}),

\begin{equation}
\begin{split}
h^2\Omega_{v}^* = 1.6 \times 10^{-1} \, v  \, \bigg( \frac{\kappa_{v}\;  \alpha}{1 + \alpha} \bigg)^2 \bigg( \frac{H}{\beta} \bigg)^1  S_{v}(f),
\\
h^2\Omega_{turb}^* = 2.01\times 10^{1} \, v \, \bigg( \frac{\kappa_{turb} \; \alpha}{1 + \alpha} \bigg)^{3/2} \bigg( \frac{H}{\beta} \bigg)^1  S_{turb}(f).
\end{split}
\end{equation}
Where unlike the runaway case, we do not assume $v = 1$ due to the bubbles reaching a terminal velocity. The MHD effieceny factor is a fraction of the sound waves, $\kappa_{turb} = \epsilon\kappa_{v}$. Current similations have motivated a range of $\epsilon \sim 0.05 - 0.10$, where we take the optimistic case of $\epsilon = 0.10$. Both contributions have unique spectral shapes, given to be,

\begin{equation}
\begin{split}
S_{v}(f) = (f/f_{p,v})^3\bigg(\frac{7}{4+3(f/f_{p,v})^2}\bigg)^{7/2},
\\
S_{turb}(f) = \frac{(f/f_{p,turb})^{3}}{(1 + f/f_{p,turb})^{11/3} (1 + 8\pi(f/H))}.
\end{split}
\end{equation}
The MHD energy density has a spectral shape dependent on the hubble rate at the time of nucleation, $H$.  Similar to the scalar spectral shape in Eq.~(\ref{eqn::spectralshape}), the frequencies are scaled by their respective temperature dependent peak frequency,

\begin{equation}
\begin{split}
f_{p,v} = 1.9 \times 10^{-5} \, \textrm{Hz}\, \frac{1}{v} \bigg( \frac{\beta}{H}\bigg)\bigg(\frac{T_{\gamma}}{100 \, \textrm{GeV}}\bigg)\bigg(\frac{g_{*}}{100}\bigg)^{\frac{1}{6}},
\\
f_{p,turb} =  2.7 \times 10^{-5} \, \textrm{Hz}\, \frac{1}{v} \bigg( \frac{\beta}{H}\bigg)\bigg(\frac{T_{\gamma}}{100 \, \textrm{GeV}}\bigg)\bigg(\frac{g_{*}}{100}\bigg)^{\frac{1}{6}},
\end{split}
\end{equation}

We evolve the frequencies and energy densities with the same redshift factors, Eq.~(\ref{eqn::redshifting}) from Sec~\ref{sec:gw}. Fig.~\ref{fig:Haasw} shows the sum of the GW energy density for the scenarios in Table~\ref{tab:cases}, but instead for the non-runaway case, with a terminal velocity of $v = 0.95$.
